---
title: "Visceral Leishmaniasis Distribution"
author: "Harry Wilson"
date: "2024-02-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This R Markdown document exhibits the code used to generate maps of visceral leishmaniasis (VL) cases both globally and surrounding the Mediterranean basin. These charts display data gathered by Pigott et al. (2014) and focus on the European region.

### Packages
You will require the following packages in order to run the code

```{r packages, warning = FALSE}
# Install Packages
library(ggplot2)
library(maps)
```


### Data
Data is taken from Pigott et al (2014) and can be downloaded [here](https://datadryad.org/stash/dataset/doi:10.5061/dryad.05f5h)

```{r map data, warning = FALSE}
# Creates a directory called "Data" within your working directory
dir.create("Data")

# Reads in csv file
VL_data <- read.csv("Data/VL_final_dataset.csv")

# Read in world coordinates from maps package
world_coordinates <- map_data("world") 
```

## European Mapping

This map displays coordinates of recorded VL cases surrounding the Mediterranean basin.

```{r maps, warning = FALSE}

ggplot() + 
  
  geom_map( data = world_coordinates, map = world_coordinates, aes(long, lat, map_id = region,), 
            color = "black", fill = "darkseagreen4", linewidth = 0.3 )+
            xlim(-11, 45)+ ylim(28, 65)+
  
  geom_map( data = VL_data, map = world_coordinates, aes(`X`, `Y`, map_id = COUNTRY,), 
            color = "black", fill = "salmon", linewidth = 0.3 )+
  
  geom_point(data = VL_data, aes(`X`, `Y`, group=YEAR, fill = YEAR), shape = 21, size=2, stroke = 0.3)+
  
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",midpoint = 1990, 
                       space = "Lab", na.value = "grey50", guide = "colourbar", aesthetics = "fill")+
  
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        panel.background = element_rect(fill = "cornflowerblue"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  
  # Titles
  annotate(geom = "label", x=4, y=64, label="Recorded Cases of Visceral Leishmaniasis",color="black", fill = "white", size = 6)+
  annotate(geom = "label", x=1.9, y=62, label="Source: Global database of leishmaniasis occurrence locations, 1960–2012",color="black", fill = "white", size = 3)

```
